#Prediction Assignment

```{r setoptions, echo=TRUE}
```
##Load and preprare the data

This code  will load the data from the working directory.  This code assumes the data has been downloaded to the working directory.


```{r}
pmltraining <- read.csv("./pml-training.csv", na.strings= c("NA",""," "))
pmltesting <- read.csv("./pml-testing.csv", na.strings= c("NA",""," "))
```

Clean the data by removing columns with NAs & identifier columns.  If we don't remove the extra data columns from the data sets then the randomForest function will not work.  It will error out.

```{r}
library(caret)
pmltrainingNAs <- apply(pmltraining, 2, function(x) {sum(is.na(x))})
pmltrainingclean <- pmltraining[,which(pmltrainingNAs == 0)]
pmltrainingclean <- pmltrainingclean[8:length(pmltrainingclean)]

pmltestingNAs <- apply(pmltesting, 2, function(x) {sum(is.na(x))})
pmltestingclean <- pmltesting[,which(pmltestingNAs == 0)]
pmltestingclean <- pmltestingclean[8:length(pmltestingclean)]
```

Create a plot to look at the data.

```{r}
featurePlot(x=pmltrainingclean[,1:4],
            y = pmltrainingclean$classe,
            plot="pairs")
```

This lattice plot shows the first 4 columns in the pmltrainingclean data set as the x values and the classe variable is the y variable.


##Build Predictors & Model

Build training and cross validation data sets

```{r}
inTrain <- createDataPartition(y = pmltrainingclean$classe, p = 0.7, list = FALSE)
training <- pmltrainingclean[inTrain, ]
crossval <- pmltrainingclean[-inTrain, ]
```

Show Summary Stats of Model

```{r}
summary(training)
```

Use the randomForest package to fit a model to predict the classe

```{r}
library(randomForest)
model <- randomForest(classe ~ ., data = training)
```

Crossvalidate the model using the confusionMatrix function to make sure model fits

```{r}
predictCrossVal <- predict(model, crossval)
confusionMatrix(crossval$classe, predictCrossVal)
```

*Kappa Statistic is 0.9936 and this reprsents the out of sample error*


Use the predect function to predict the classes of the testing data set

```{r}
predictTest <- predict(model, pmltestingclean)
predictTest
```
